SIMD-vectorized Chemical Source Term Evaluation

### 2017-02-25 paper.tex
clarify loading examples with Nv

\begin{figure}[htb]
  \centering
  \begin{minipage}[t]{0.45\linewidth}
      \includegraphics[width=\textwidth]{SIMD.pdf}
      \caption{Schematic of SIMD processing.  A single processing element (e.g. a CPU core) contains a vector unit with $N$ lanes ($\text{L}_0,\ldots \text{L}_N$).  The vector unit executes a single instruction concurrently on multiple data.}
      \label{F:SIMD}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.45\linewidth}
      \includegraphics[width=\textwidth]{mem_access.pdf}
      \caption{Representation of an 2-D array with dimension $N \times K$ in CPU memory, with linear index marked.}
      \label{F:mem}
  \end{minipage}
\end{figure}

@@

In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $1, 2\ldots$ will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state $j$, thus SIMD-lanes loading species concentrations $k, k+1\ldots$ for e.g., state 1, will access sequential memory locations if the data is stored in ``C''-order.
Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
However, a deep vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g. from differing for-loop bounds).

>>>>

\begin{figure}[htb]
  \centering
  \begin{minipage}[t]{0.45\linewidth}
      \includegraphics[width=\textwidth]{SIMD.pdf}
      \caption{Schematic of SIMD processing.  A single processing element (e.g. a CPU core) contains a vector unit with $N_v$ lanes ($\text{L}_0,\ldots \text{L}_{N_v}$).  The vector unit executes a single instruction concurrently on multiple data.}
      \label{F:SIMD}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.45\linewidth}
      \includegraphics[width=\textwidth]{mem_access.pdf}
      \caption{Representation of an 2-D array with dimension $N \times K$ in CPU memory, with linear index marked.}
      \label{F:mem}
  \end{minipage}
\end{figure}

@@

In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for e.g., states $1, 2,\ldots N_{v}$---the number of vector lanes---will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k, k+1\ldots k + N_v$ for e.g., state 1, will access sequential memory locations if the data is stored in ``C''-order.
Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
However, a deep vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g. from differing for-loop bounds).
